---
title: "coursework7"
author: "Fortunat Mutunda"
date: "March 23, 2016"
output: pdf_document
---
```{r, echo=FALSE,message=FALSE,warning=FALSE}

library(arules)
library(arulesViz)
library(ggplot2)
```
1. Report which tools you decided to use, how you used them, what were the first results. Also report the running times for the tools chosen.

***comment***
*Mine rules from the supermarket txt*
1. Run whatever tool you run and plot the running time 
2. Get suppeort and confidence
3. Which tool did you use , how did you use it how long did it run.
4. Discrive in general... show examples, max, min,... etc.
5. How did you get the result what params etc...

```{r}
#supermarket = read.csv("supermarket.txt", header = F, sep = " ")
#write.csv(supermarket, file = "supermarket.csv")
trans = read.transactions("supermarket.txt", format = "basket" , sep = " ")
#basket = read.transactions("supermarket.csv", format = "basket" , sep = ",")
ruless <- apriori(trans, parameter= list(supp=0.02))

image(trans)
itemFrequencyPlot(trans, support = 0.02)
plot(ruless, method = "graph")
plot(ruless, method="paracoord", control=list(reorder=TRUE))

```

2. Report overall 10 different high-support, high-confidence, high-lift rules; provide the respective contingency tables and scores.
***contengancy table ***
if I can find a tool that get me a contengancy table.
show the top 10 rules.

3. Discuss whether some other scores studied last week or in the lecture slides would help identify "more interesting" and different rules?
***Use something from last homewok***
no confidence or lift or anything try other maesures 

4. Given the ability to discover frequent itemsets and association rules, propose a strategy to use these tools to study different customer segments, shops, shopping times, or specific products.
***What can you do with the data?***
Just write something, what method toools you have learnt so far to study diff rules...
Suggest a plan...etc.
There is nothing specific, just be creative... In which part can you use frequent set item or item set mining 

5. Select some relatively high-support high-confidence rule (A->B) and based on that example describe the conditional probabilities P(A|B) and P(B|A), as well as the Bayes rule.

*** Just learn what a conditional probalities ***

6. (Bonus 2p) Run Krimp on same data, provide commands and describe your findings and compare to Frequent Itemset Mining + Association rules. (link to Krimp documentation)

***Krimp ***





```{r, echo=FALSE}

```


